On the presentation of orthogonal groups over polynomial rings

### 2019-04-29 dl-paper.tex
Finished rewriting 4.1

\subsection{Computation of the kernel of the map of evaluation at $0$.}
%TODO: Do we need any specific conditions on \Phi? Specify them.
For a pair $(R, I)$ we denote by $\overline{\St}(\Phi, R, I)$ the image of the map $\mu$ from~\eqref{relSteinberg-def}, in particular, it is a normal subgroup of $\St(\Phi, R)$.
We also denote by $\St(\Phi, I)$ the subgroup of $\St(\Phi, R)$ generated as a group by root unipotents of level $I$.
It is clear that $\overline{\St}(\Phi, R, I)$ contains $\St(\Phi, I)$ and, in fact, is its normal closure.

The aim of this subsection is to describe the kernel of the map $ev_{X=0}^*\colon\overline{\St}(\Phi, A[X], M[X]) \to \overline{\St}(\Phi, A, M)$
induced by the ring homomorphism of evaluation at $0$. Clearly, this kernel, denote it by $K$, contains the subgroup $\overline{\St}(\Phi, A[X], XM[X])$, however, apriori it may be larger,
 so our goal will be to describe the explicit extra elements sufficient to generate it.

Obviously, $ev_{X=0}^*$ admits a section hence we can consider $\overline{\St}(\Phi, A, M)$ as a subgroup of $\overline{\St}(\Phi, A[X], M[X])$,
 moreover, the following decomposition holds:
\begin{equation} \label{relZero-decomp} \overline{\St}(\Phi, A[X], M[X]) = \overline{\St}(\Phi, A, M) \cdot K.\end{equation}
\begin{lemma} 
The following decomposition holds
 \[ K = \overline{\St}(\Phi, A[X], XM[X]) \cdot \left[\St(\Phi, XA[X]),\ \overline{\St}(\Phi, A, M)\right].\] 
\end{lemma}
\begin{proof}
 Fix $g \in K$ and write it as $g(X) = \prod_i z_{\alpha_i}(f_i(X), \xi_i(X))$ for some $f_i(X) = f_i(0) + Xf_i'(X) \in M[X]$, $\xi_i(X) = \xi_i(0) + X\xi_i'(X) \in A[X]$.
 It is clear that modulo $\overline{\St}(\Phi, A[X], XM[X])$ the element $g(X)$ is congruent to $g_1(X) = \prod_i z_{\alpha_i}(f_i(0), \xi_i(X)).$

 Now each factor $z_{\alpha_i}(f_i(0), \xi_i(X))$ can be written as follows:
 \[z_{\alpha_i}(f_i(0), \xi_i(0))^{x_{-\alpha_i}(X\xi'_i(X))} = [x_{-\alpha_i}(X\xi'_i(X)), z_{\alpha_i}(f_i(0), \xi_i(0))] \cdot z_{\alpha_i}(f_i(0), \xi_i(0)).\]
 Since the subgroup $C = \left[\St(\Phi, XA[X]), \overline{\St}(\Phi, A, M)\right]$ is normalized by $\overline{\St}(\Phi, A, M)$ 
 (this follows e.\,g. from the formula $[g, h]^{h_1} = [h_1^{-1}, g][g, h_1^{-1}h]$) we conclude that $g_1(X)$ is congruent to $\prod_i z_{\alpha_i}(f_i(0), \xi_i(0)) = g(0) = 1$ modulo $C$,
 which implies the assertion. \qedhere

 %Denote by $H$ the composite of $C$ and $\overline{\St}(\Phi, A[X], XM[X])$. It is clear that $H \subseteq I$. The reverse inclusion follows from the above argument and the fact that for $g \in I$ one has $g(0)=1$. \qedhere 
\end{proof}
 \begin{lemma} If $A$ is a local ring with maximal ideal $M$ then \[\overline{\St}(\Phi, A, M) = U(\Phi^+, M) \cdot M(\Phi, A, M) \cdot U(\Phi^-, M).\] \end{lemma}
 \begin{proof} See~\cite[Theorem~2.4]{Ste73}. \end{proof}

\begin{corollary} For $A$, $M$ as above holds
 \[\left[\St(\Phi, XA[X]), \overline{\St}(\Phi, A, M)\right] \subseteq \St(\Phi, XM[X]) \cdot \left[\St(\Phi, XA[X]), \St(\Phi, M)\right].\]
\end{corollary}
\begin{proof}
 Notice first that the generators of $M(\Phi, A, M)$ normalize both $\St(\Phi, XA[X])$ and $\St(\Phi, M)$, see~\eqref{eq:conj-h-x}.
 Now for $g \in \St(\Phi, XR[X])$, $h \in M(\Phi, A, M)$, $u^+ \in U(\Phi^+, M)$, $u^- \in U(\Phi^-, M)$ holds:
 \[ [g, h u^+ u^-] = [g, h] \cdot [{}^{h}\!g, {}^{h}\!(u^+u^-)] \in \St(\Phi, XM[X]) \cdot [\St(\Phi, XA[X]), \St(\Phi, M)]. \qedhere\]
\end{proof}

\begin{corollary} For $A$ and $M$ as above the subgroup $K$ is generated as a group by its subgroup $\St(\Phi, A[X], XM[X])$ and
 the elements $[x_\alpha(m), x_{-\alpha}(X\xi)]$, $m \in M$, $\xi \in A[X]$, $\alpha \in \Phi$. 
\end{corollary}
\begin{proof}
By the above lemmata one has $K \subseteq \St(\Phi, A[X], XM[X]) \cdot [\St(\Phi, XA[X]), \St(\Phi, M)].$
It is clear that modulo $\St(\Phi, A[X], XM[X])$ the commutator subgroup $[\St(\Phi, XA[X]), \St(\Phi, M)]$ is generated
 by all $[x_\alpha(m), x_{-\alpha}(X\xi)]^g$, $m \in M$, $\xi \in A[X]$, where $g \in \St(\Phi, M + XA[X])$.
To verify the assertion, therefore, it remains to show that the commutators $[[x_\alpha(m), x_{-\alpha}(X\xi)], g]$ belong to $\St(\Phi, A[X], XM[X])$.
Moreover, since the latter subgroup is normal it suffices to prove this inclusion in the special case when $g$ is a generator of $\St(\Phi, A[X])$,
i.\,e. $g = x_\beta(\xi)$, $\eta \in A[X]$. Without loss of generality we may assume $\alpha \neq \pm \beta$.
\end{proof}

>>>>

\subsection{Computation of the kernel of the map of evaluation at $0$.}
For a pair $(R, I)$ we denote by $\overline{\St}(\Phi, R, I)$ the image of the map $\mu$ from~\eqref{relSteinberg-def}, in particular, it is a normal subgroup of $\St(\Phi, R)$.
We also denote by $\St(\Phi, I)$ the subgroup of $\St(\Phi, R)$ generated as a group by root unipotents of level $I$.
It is clear that $\overline{\St}(\Phi, R, I)$ contains $\St(\Phi, I)$ and, in fact, is its normal closure.
We also denote by $\overline{H}(\Phi, R, I)$ the subgroup of $\overline{\St}(\Phi, R, I)$ generated by the semisimple root elements $h_\alpha(u)$ and symbols $\{u, v\}$, $u \in (1+I)^*$, $v \in R^*$, $\alpha\in \Phi$.

The aim of this subsection is to describe the kernel of the map $ev_{X=0}^*\colon\overline{\St}(\Phi, A[X], M[X]) \to \overline{\St}(\Phi, A, M)$
induced by the ring homomorphism of evaluation at $0$. Clearly, this kernel, denote it by $K(A[X], M[X])$, contains the subgroup $\overline{\St}(\Phi, A[X], XM[X])$, however, apriori it may be larger, so our goal will be to describe the explicit extra elements sufficient to generate it.

Obviously, $ev_{X=0}^*$ admits a section hence we can consider $\overline{\St}(\Phi, A, M)$ and $\St(\Phi, M)$ as subgroups of $\overline{\St}(\Phi, A[X], M[X])$,
 moreover, the following decomposition holds:
\begin{equation} \label{relZero-decomp} \overline{\St}(\Phi, A[X], M[X]) = \overline{\St}(\Phi, A, M) \cdot K(A[X], M[X]).\end{equation}
\begin{lemma} 
The following decomposition holds
 \[ K(A[X], M[X]) = \overline{\St}(\Phi, A[X], XM[X]) \cdot \left[\St(\Phi, XA[X]),\ \overline{\St}(\Phi, A, M)\right].\] 
\end{lemma}
\begin{proof}
 Fix $g \in K(A[X], M[X])$ and write it as $g(X) = \prod_i z_{\alpha_i}(f_i(X), \xi_i(X))$ for some $f_i(X) = f_i(0) + Xf_i'(X) \in M[X]$, $\xi_i(X) = \xi_i(0) + X\xi_i'(X) \in A[X]$.
 It is clear that modulo $\overline{\St}(\Phi, A[X], XM[X])$ the element $g(X)$ is congruent to $g_1(X) = \prod_i z_{\alpha_i}(f_i(0), \xi_i(X)).$

 Now each factor $z_{\alpha_i}(f_i(0), \xi_i(X))$ can be written as follows:
 \[z_{\alpha_i}(f_i(0), \xi_i(0))^{x_{-\alpha_i}(X\xi'_i(X))} = [x_{-\alpha_i}(X\xi'_i(X)), z_{\alpha_i}(f_i(0), \xi_i(0))] \cdot z_{\alpha_i}(f_i(0), \xi_i(0)).\]
 Since the subgroup $C = \left[\St(\Phi, XA[X]), \overline{\St}(\Phi, A, M)\right]$ is normalized by $\overline{\St}(\Phi, A, M)$ 
 (this follows e.\,g. from the formula $[g, h]^{h_1} = [h_1^{-1}, g][g, h_1^{-1}h]$) we conclude that $g_1(X)$ is congruent to $\prod_i z_{\alpha_i}(f_i(0), \xi_i(0)) = g(0) = 1$ modulo $C$,
 which implies the assertion. \qedhere
\end{proof}

 \begin{theorem}[Stein] \label{thmStein} For a local ring $A$ with maximal ideal $M$ holds \[\overline{\St}(\Phi, A, M) = U(\Phi^+, M) \cdot \overline{H}(\Phi, A, M) \cdot U(\Phi^-, M).\] \end{theorem} \begin{proof} See~\cite[Theorem~2.4]{Ste73}. \end{proof}

\begin{prop} \label{Kgen}
For $A$ and $M$ as above the subgroup $K(A[X], M[X])$ is generated as a group by its subgroup $\St(\Phi, A[X], XM[X])$ and
 the elements $[x_\alpha(m), x_{-\alpha}(X\xi)]$, $m \in M$, $\xi \in A[X]$, $\alpha \in \Phi$. \end{prop}
\begin{proof}
 Notice that it follows from~\eqref{eq:conj-h-x} $\overline{H}(\Phi, A, M)$ normalizes both $\St(\Phi, XA[X])$ and $\St(\Phi, M)$ and, moreover, that $[\overline{H}(\Phi, A, M),\ \St(\Phi, XA[X])] \subseteq \St(\Phi, XM[X])$.

Now for $g \in \St(\Phi, XA[X])$, $h \in \overline{H}(\Phi, A, M)$, $u^+ \in U(\Phi^+, M)$, $u^- \in U(\Phi^-, M)$ holds:
\[ [g, h u^+ u^-] = [g, h] \cdot [{}^{h}\!g, {}^{h}\!(u^+u^-)] \in \St(\Phi, XM[X]) \cdot [\St(\Phi, XA[X]), \St(\Phi, M)],\]
hence $\left[\St(\Phi, XA[X]), \overline{\St}(\Phi, A, M)\right] \subseteq \St(\Phi, XM[X]) \cdot \left[\St(\Phi, XA[X]), \St(\Phi, M)\right]$ and therefore \[K(A[X], M[X]) = \overline{\St}(\Phi, A[X], XM[X]) \cdot \left[\St(\Phi, XA[X]),\ \St(\Phi, M)\right].\]

It is clear that modulo $\St(\Phi, A[X], XM[X])$ the commutator subgroup $[\St(\Phi, XA[X]), \St(\Phi, M)]$ is generated by elements of the form $[x_\alpha(m), x_{-\alpha}(X\xi)]^g$, where $m \in M$, $\xi \in A[X]$, $g \in \St(\Phi, A[X])$.
Thus, to verify the assertion it remains to show that the commutators $[[x_\alpha(m), x_{-\alpha}(X\xi)], g]$ belong to $\St(\Phi, A[X], XM[X])$.
Since the latter subgroup is normal it suffices to prove this inclusion in the special case when $g$ is a member of some generating set for $\St(\Phi, A[X])$.
Clearly, the set consisting of $x_\beta(\xi)$, $\xi \in A[X]$, $\beta \neq \pm \alpha$ is such a generating set, but in this case the required inclusions follow from~\eqref{C1}--\eqref{C3} of~\cref{Crels}. \end{proof}

\begin{corollary} \label{Kgen-strong}
 For $A$ and $M$ as in~\cref{thmStein} and arbitrary fixed root $\gamma\in \Phi$ the subgroup $K(A[X], M[X])$ is generated as a group by $\St(\Phi, A[X], XM[X])$ and the elements $c_{\pm \gamma}(f, \eta)$, $f \in M[X]$, $\eta \in XA[X]$.
\end{corollary}
\begin{proof}
 Choose some ordering on $\Phi$ with respect to which $\gamma$ becomes the maximal positive root. Invoking~\eqref{C4} of~\cref{Crels} with $I = M[X]$, $J=XA[X]$, $\xi = 1$ we obtain that modulo $\overline{\St}(\Phi, A[X], XM[X])$ the element $c_{\alpha + \beta}(f, \eta)$ is equivalent to $c_{\alpha}(-\epsilon f, -\epsilon \eta)^{-1}$, $\epsilon = N_{\alpha, \beta}$. Arguing by induction on the height of $\alpha$ we conclude that all $c_\alpha(f, \eta)$, $\alpha \in \Phi^+$ belong to the subgroup generated by $\overline{\St}(\Phi, A[X], XA[X])$ and $c_\gamma(f, \eta)$. Similar argument also works for the negative roots. The required assertion now follows from the above proposition.
\end{proof}
